General constraints on magic distillation protocols from statistical mechanics

### 2021-06-01 magic_paper.tex
added refs and lemma on additivity of H. Will continue shortly

\section{Introduction}
\label{sec:intro}
The past few years have seen rapid progress towards the goal of a fault-tolerant quantum computer~\cite{Nickerson_2014, Nikahd_2017, chao_2018, lin_pieceable_2020, Lin_2020, Bourassa_2021}. However, many challenges remain and there is increasing need for theory to take into account physical limitations of the hardware involved. The surface code~\cite{Bravyi_1998, Freedman_2001, Dennis_2002, Raussendorf_2007} is a leading framework for fault-tolerance with very high error thresholds. Within this scheme, Clifford unitaries can be implemented in a robust, fault-tolerant way. However, due to the Eastin-Knill theorem~\cite{Eastin_2009}, we also know that it is impossible to have a universal set of transversal gates, and although Clifford unitaries can be realised transversally~\cite{Calderbank_1996, Steane_1996}, one needs to find ways around the Eastin-Knill restriction. This can be achieved by injecting in quantum states, called magic states, which promote the Clifford group to universal quantum computing~\cite{cit:bravyi}. The obstacle to this is that these states are invariably noisy and so protocols involving stabilizer operations must be employed to purify many copies of the magic states and improve the overall performance of the induced quantum gates. A central question then arises about the overhead on purifying many copies of a magic state into less noisy forms.

@@

\newpage
\subsection{Single-shot divergences for quasi-distributions}
As already mentioned, the Wigner distribution is generally a quasi-distribution and so taking the Shannon entropy $H(\p) = -\sum_i p_i \log p_i$ of the Wigner distribution is not possible off of the set $\F$. However, does the same obstacle applies to Renyi entropies? The $\alpha$--Renyi entropy $H_\alpha(\p)$ is defined as
\begin{equation}
H_\alpha(\p) := -\frac{1}{\alpha -1} \log \sum_i p_i^\alpha,
\end{equation}
where $\alpha \ge 0$ and $\lim_{\alpha \rightarrow 1} H_\alpha(\p) = H(\p)$. It is clear that if we take $\alpha$ to be an even integer then $H_\alpha (W_\rho(\x))$ is always a well-defined function on Wigner quasi-distributions. However it still remains to link such expressions to majorization of quasi-distributions.

In this direction we first note that the analysis for the temperature-dependent bound given by Equation (\ref{eq:rate_bounds_proof}) implicitly makes use of the single-shot Renyi divergence $D_\infty (\p ||\r)$, which is given by
\begin{equation}
D_\infty (\p ||\r) = \log \max_i \frac{p_i}{r_i}.
\end{equation}
This motivates defining the following \emph{Wigner divergence} $D_\infty(\rho, \tau) := D_\infty (W_\rho(\x) || W_\tau (\x))$ on all quantum states. In Appendix~\ref{app:lcst_technical}, we show that this divergence has the following natural properties that are derived via relative majorization.
\begin{theorem} Let $\tau$ be in the interior of $\F$, and define the Wigner divergence $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$.
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
See Appendix~\ref{app:lcst_technical} for the proof of this. Moreover, in the appendices we also prove that for unital processes the Renyi entropies $H_{2m}(\w)$ are monotones for any $m \in \mathbb{N}$ and lead the generalized distillation bounds.
\begin{equation}
R \le \frac{H_{2m}(\rho)}{H_{2m}(\rho')}.
\end{equation}
It is of interest to further explore this perspective and determine what components of single-shot information theory could be extended to quasi-distributions.

@@

\begin{theorem} Let $\tau$ be in the interior of $\F$, and define $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$.
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
\begin{proof}Since $\tau$ is in the interior of $\F$ its Wigner function obeys $W_\tau(\x) >0$ for all points $\x$ in the phase space. In general $W_\rho(\x)$ is a quasi-distribution, however since $\sum_{\x} W_\rho (\x) = 1$ this implies it must have at least one positive component and so $\max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)}>0$. Therefore $D_\infty (\rho,\tau)$ is always well-defined.

@@

\begin{definition} A continuous, real-valued function $f$ defined on $\D \subseteq \mathbb{R}^N$ is called Schur-convex on the domain $\D$ if $\p \prec \q$ on $\D$ implies that $f(\p) \le f(\q)$. A continuous function $f:\D \rightarrow \mathbb{R}$ is called Schur-concave if $(-f)$ is Schur-convex.
\end{definition}
We now can consider the set of all quasi-distributions, which form a hyper-plane and are given by
\begin{equation}
\D = \{ \w \in \mathbb{R}^N: \sum_i w_i = 1\}.
\end{equation}
For functions that are symmetric, namely $f(\p) = f(\Pi \p)$ for any permutation $\Pi$ of the components, it suffices to restrict to the following region,

@@

We now have the following theorem [CITE Marshall+Olkin] that provides a characterization of symmetric, Schur-convex functions on quasi-distributions.
\begin{theorem} Let $f$ be a real-valued continuous function defined on $\D^\downarrow$ that is continuously differentiable on the interior of $\D^\downarrow$. Then $f$ is Schur-convex on $\D^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \D^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\D^\downarrow$.
\end{theorem}

\ddd{[TO-DO AND TIDY AND CHECK]} Okay great...this looks easy to do, we just need to compute the terms
\begin{equation}
-\frac{\partial H_\alpha}{\partial w_k} (w_1, w_2, \dots, w_N),
\end{equation}
for Renyi entropies (which are symmetric), and look to see when the above inequality holds. 
We assume $ \alpha=2m$ is an even integer, so that $H_\alpha (\w)$ is well-defined on quasi-distributions. We then have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{2m} \\
&= \frac{1}{(2m-1) \sum_i w_i^{2m}}2mw_k^{2m-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^{2m}$ for simplicity, that
\begin{align}
 \frac{1}{(2m-1)S}2mw_i^{2m-1}\ge \frac{1}{(2m-1)S}2mw_{i+1}^{2m-1},
\end{align}
for each $i=1,2,\dots, N-1$. We assume that $m >0$ and $2m \ne 1$ and so this becomes
\begin{equation}
w_i^{2m-1}\ge w_{i+1}^{2m-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $2m >1$ this condition holds whenever $w_i \ge w_{i+1}$, which is true in $\D^\downarrow$. Therefore $-H_{2m}(\w)$ is Schur-Convex on quasi-distributions for any integer $m\in \mathbb{N}$. This implies it is a monotone function for majorization, namely,
\begin{equation}
\w \succ \w' \mbox{ implies } H_{2m} (\w) \le H_{2m} (\w') \mbox{for any } m \in \mathbb{N}.
\end{equation}
This therefore should give \emph{additional} distillation bounds in the unital fragment. We'd need to check additivity is okay, but this should be fine. Therefore
\begin{equation}
W_{\rho_n}(\x) \succ W_{\rho'_m}(\x) \mbox{ implies } n H_{2m}(\rho) \le m H_{2m}(\rho'). 
\end{equation}
(Some sign error here? Should get:)
\begin{equation}
R \le \frac{H_{2m}(\rho)}{H_{2m}(\rho')}.
\end{equation}
\ddd{[Plan: check all these calculations! Give the generalised unital bound. Then we just conjecture this extends to Renyi-divergences and leave it at this.]}
\subsection{Dropping pre/post-processing of the magic states}
\label{app:cliff_processing}

>>>>

\section{Introduction}
\label{sec:intro}

\nick{Bunch of refs:}~\cite{White_2021, Rethinasamy_2020, Blackwell_1953, Sarkar_2020, Gour_2020}

The past few years have seen rapid progress towards the goal of a fault-tolerant quantum computer~\cite{Nickerson_2014, Nikahd_2017, chao_2018, lin_pieceable_2020, Lin_2020, Bourassa_2021}. However, many challenges remain and there is increasing need for theory to take into account physical limitations of the hardware involved. The surface code~\cite{Bravyi_1998, Freedman_2001, Dennis_2002, Raussendorf_2007} is a leading framework for fault-tolerance with very high error thresholds. Within this scheme, Clifford unitaries can be implemented in a robust, fault-tolerant way. However, due to the Eastin-Knill theorem~\cite{Eastin_2009}, we also know that it is impossible to have a universal set of transversal gates, and although Clifford unitaries can be realised transversally~\cite{Calderbank_1996, Steane_1996}, one needs to find ways around the Eastin-Knill restriction. This can be achieved by injecting in quantum states, called magic states, which promote the Clifford group to universal quantum computing~\cite{cit:bravyi}. The obstacle to this is that these states are invariably noisy and so protocols involving stabilizer operations must be employed to purify many copies of the magic states and improve the overall performance of the induced quantum gates. A central question then arises about the overhead on purifying many copies of a magic state into less noisy forms.

@@

\newpage
\subsection{Single-shot divergences for quasi-distributions}
As already mentioned, the Wigner distribution is generally a quasi-distribution and so taking the Shannon entropy $H(\p) = -\sum_i p_i \log p_i$ of the Wigner distribution is not possible off of the set $\F$. However, does the same obstacle apply to Renyi entropies? The $\alpha$--Renyi entropy $H_\alpha(\p)$ is defined as
\begin{equation}
	H_\alpha(\p) := \frac{1}{1-\alpha} \log \sum_i p_i^\alpha,
\end{equation}
where $\alpha \ge 0$ and $\lim_{\alpha \rightarrow 1} H_\alpha(\p) = H(\p)$. It is clear that if we take $\alpha = 2m$ to be an even integer then $H_{2m}(\rho) := H_{2m}(W_\rho(\x))$ is always a well-defined function on Wigner quasi-distributions. However, it still remains to link such expressions to majorization of quasi-distributions.

In this direction, we first note that the analysis for the temperature-dependent bound given by Equation (\ref{eq:rate_bounds_proof}) implicitly makes use of the single-shot Renyi divergence $D_\infty (\p ||\r)$, which is given by
\begin{equation}
D_\infty (\p ||\r) = \log \max_i \frac{p_i}{r_i}.
\end{equation}
This motivates defining the following \emph{Wigner divergence} $D_\infty(\rho, \tau) := D_\infty (W_\rho(\x) || W_\tau (\x))$ on all quantum states. In Appendix~\ref{app:lcst_technical}, we show that this divergence has the following natural properties that are derived via relative majorization.
\begin{theorem} Let $\tau$ be in the interior of $\F$, and define the Wigner divergence $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$.
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
See Appendix~\ref{app:lcst_technical} for the proof of this. Moreover, in the appendices we also prove that for unital processes the Renyi entropies $H_{2m}(\w)$ are monotones for any $m \in \mathbb{N}$ and lead the generalized distillation bounds.
\begin{equation}
R \le \frac{H_{2m}(\rho)}{H_{2m}(\rho')}.
\end{equation}
It is of interest to further explore this perspective and determine what components of single-shot information theory could be extended to quasi-distributions.

@@

\begin{theorem} Let $\tau$ be in the interior of $\F$, and define $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$.
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
\begin{proof}
	Since $\tau$ is in the interior of $\F$, its Wigner function obeys $W_\tau(\x) >0$ for all points $\x$ in the phase space. In general, $W_\rho(\x)$ is a quasi-distribution, however since $\sum_{\x} W_\rho (\x) = 1$ this implies it must have at least one positive component, so $\max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)}>0$. Therefore $D_\infty (\rho,\tau)$ is always well-defined.

@@

\begin{definition} A continuous, real-valued function $f$ defined on $\D \subseteq \mathbb{R}^N$ is called Schur-convex on the domain $\D$ if $\p \prec \q$ on $\D$ implies that $f(\p) \le f(\q)$. A continuous function $f:\D \rightarrow \mathbb{R}$ is called Schur-concave if $(-f)$ is Schur-convex.
\end{definition}
We now can consider the set of all quasi-distributions, which form a hyper-plane and are given by
\begin{equation}
\D = \{ \w \in \mathbb{R}^N: \sum_i w_i = 1\}.
\end{equation}
For functions that are symmetric, namely $f(\p) = f(\Pi \p)$ for any component permutation $\Pi$, it suffices to restrict to the following region,

@@

We now have the following theorem~\cite{cit:marshall} that provides a characterization of symmetric, Schur-convex functions on quasi-distributions.
\begin{theorem} Let $f$ be a real-valued continuous function defined on $\D^\downarrow$ that is continuously differentiable on the interior of $\D^\downarrow$. Then $f$ is Schur-convex on $\D^\downarrow$ if and only if $\partial_k f(\w)$ is decreasing in $k$ for all $\w \in \D^\downarrow$, namely
\begin{equation}
\partial_1 f(\w) \ge \partial_2 f(\w) \ge \cdots \ge \partial_N f(\w),
\end{equation}
for all $\w$ in the interior of $\D^\downarrow$.
\end{theorem}

\begin{lemma}[Additivity of $H_{2m}$]\label{H_add}
	For any states $\rho, \rho'$ and non-negative $m$, we have
	\begin{equation}
		H_{2m}(\rho \otimes \rho') = H_{2m}(\rho) + H_{2m}(\rho')
	\end{equation}
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{2m}&(\rho \otimes \rho') = \frac{1}{1-2m}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{2m} \\
		&= \frac{1}{1-2m}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{2m} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{2m} \right) \\
		&= H_{2m}(\rho) + H_{2m}(\rho').
	\end{align*}
\end{proof}
\cref{H_add} implies that for any $n$,
\begin{equation}
	H_{2m}(\rho_n) := H_{2m}(\rho^{\otimes n}) = n H_{2m}(\rho).
\end{equation}

Using the above results, we can show that $H_{2m}$ is Schur-concave for all quasi-distributions.
In particular, for any non-negative integer $m$, we have that
\begin{equation}
	W_{\rho_n}(\bmx) \succ W_{\rho'_{n'}}(\bmx) \mbox{ implies } H_{2m}(\rho_n) \leq H_{2m}(\rho'_{n'}).
\end{equation}
By additivity of $H_{2m}$, we get $n H_{2m}(\rho) \leq n' H_{2m}(\rho')$, or equivalently,
\begin{equation}
	\frac{n'}{n} \geq \frac{H_{2m}(\rho)}{H_{2m}(\rho')}.
\end{equation}

\nick{Wrong way? I don't think that $m \rightarrow \infty$ leads to our bound}

\ddd{[TO-DO AND TIDY AND CHECK]} Okay great...this looks easy to do, we just need to compute the terms
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_k} (w_1, w_2, \dots, w_N),
\end{equation}
for Renyi entropies (which are symmetric), and look to see when the above inequality holds. 
We assume $ \alpha=2m$ is an even integer, so that $H_\alpha (\w)$ is well-defined on quasi-distributions. We then have that
\begin{align}
V_k:=-\frac{\partial H_\alpha}{\partial w_k} &= \partial_k \frac{1}{\alpha-1} \log \sum_i w_i^{2m} \\
&= \frac{1}{(2m-1) \sum_i w_i^{2m}}2mw_k^{2m-1}.
\end{align}
We now impose the monotonicity condition $V_1 \ge V_2 \ge V_3 \cdots \ge V_N$. This requires, setting $S= \sum_i w_i^{2m}$ for simplicity, that
\begin{align}
 \frac{1}{(2m-1)S}2mw_i^{2m-1}\ge \frac{1}{(2m-1)S}2mw_{i+1}^{2m-1},
\end{align}
for each $i=1,2,\dots, N-1$. We assume that $m >0$ and $2m \ne 1$ and so this becomes
\begin{equation}
w_i^{2m-1}\ge w_{i+1}^{2m-1}\mbox{ for all } 1 \le i < N.
\end{equation}
For $2m >1$ this condition holds whenever $w_i \ge w_{i+1}$, which is true in $\D^\downarrow$. Therefore $-H_{2m}(\w)$ is Schur-Convex on quasi-distributions for any integer $m\in \mathbb{N}$. This implies it is a monotone function for majorization, namely,
\begin{equation}
\w \succ \w' \mbox{ implies } H_{2m} (\w) \le H_{2m} (\w') \mbox{for any } m \in \mathbb{N}.
\end{equation}
This therefore should give \emph{additional} distillation bounds in the unital fragment. We'd need to check additivity is okay, but this should be fine. Therefore
\begin{equation}
W_{\rho_n}(\x) \succ W_{\rho'_m}(\x) \mbox{ implies } n H_{2m}(\rho) \le m H_{2m}(\rho'). 
\end{equation}
(Some sign error here? Should get:)
\begin{equation}
R \le \frac{H_{2m}(\rho)}{H_{2m}(\rho')}.
\end{equation}
\ddd{[Plan: check all these calculations! Give the generalised unital bound. Then we just conjecture this extends to Renyi-divergences and leave it at this.]}
\subsection{Dropping pre/post-processing of the magic states}
\label{app:cliff_processing}
